Recursive Neural Network Part-of-Speech Tagger

### 2013-12-21 paper/paper.tex
Minor fixes.

For instance, given the sentence $(w_{i_1},w_{i_2},\ldots,w_{i_K})$ of length $K$, we first calculate the word representations $V\Ket{w_{i_1}},\ldots,V\Ket{w_{i_K}}$, then we calculate each phrase representation $p_j$ from the previous phrase representation $p_{j-1}$ and the current word representation $V\Ket{w_{i_j}}$ by concatenating the two representation vectors, multiplying by $W_{combine}$ and adding $b_{combine}$ before applying the $tanh(\cdot)$ function:

$p_j = tanh(W_{combine}\Big[p_{j-1};V\Ket{w_{i_j}}\Big] + b_{combine})$

Subsequently, each $p_j$ is passed along to the classifier, which applies the operation:

$t_{i_j} = softmax(W_{classify}p_j + b_{classify})$

to obtain the predicted tag $t_{i_j}$ for $w_{i_j}$.

@@

Training neural networks can be tricky since the backpropagation algorithm relies upon a learning rate parameter as well as initial weight and bias values. The learning rate should not be too large, otherwise the gradient descent will skip over local or global minima. However, it should also not be too small, otherwise training may take too long. It perhaps should vary over the course of training and decrease in order to better fine tune the fit to the minima. However, it's not obvious exactly which set of choices will be best.

@@

\subsubsection{AdaGrad}
We used the AdaGrad online learning method to train our model (Socher et al., 2013). Let $\theta$ be a vector of all parameters needed to train an RNN model and $\theta_t$ means the parameter vector before training on the $t$-th sentence. And let  $g_{t,i}$ to be the gradient value of $i_{th}$ parameter on $t$-th sentence. Then the AdaGrad is just updating the parameter vector by
$$\theta_{{t+1},i} = \theta_{{t},i}  - \frac{\alpha}{\sqrt{\sum_{s=1}^t \theta_{{s},i}^2 }} g_{t,i}$$,
where the $\alpha$ is a constance to set the scale of learning rate.

We compared the AdaGrad with fixed learning rate online traning. The Figure \ref{Adagrad_fig} is the accuarcy curve on training iteration. From Figure \ref{Adagrad_fig}, it's quite clear that the AdaGrad converges faster and better than fixed learning rate.

@@

\begin{itemize}
\item \textbf{Same word test}:  We chose some words with more than one common tag and saw whether our model predicted those tags correctly. We tested on `to' (possible tags are TO and IN) and `work' (common possible tags are NN, VB and VBP). Our RNN predicts `to' with 0.80 in-domain and 0.77 out-of-domain accuracy, covering TO and IN. The RNN predicts `work' with 0.88 in-domain and 0.83 out-of-domain accuracy, covering NN, VB and VBP.
\item \textbf{Disconnected core module test}: We disconnected the sequential information passing by ignoring the input embeddings from lower layers, and backward gradient from upper layers. So although it's still an RNN, there is actually no information passed between layers. We trained this non-sequential RNN and the original RNN with 100 sentences and 100 iterations, same random seed and training parameters. The Figure \ref{Noseq_fig} shows the out-of-domain accuracy of two models by the number of iterations, which can be also seen from Figure \ref{Leftsize_fig}. The difference is not small. So we think the sequential information was passed and used at least partially correctly.
\end{itemize}
\begin{figure}
\includegraphics[scale=0.5]{outdomain_noseq.png}
\caption{Comparison of non-sequential RNN and original RNN for disconnected core module test}\label{Noseq_fig}
\end{figure}

>>>>

For instance, given the sentence $(w_{i_1},w_{i_2},\ldots,w_{i_K})$ of length $K$, we first calculate the word representations $V\Ket{w_{i_1}},\ldots,V\Ket{w_{i_K}}$, then we calculate each phrase representation $p_j$ from the previous phrase representation $p_{j-1}$ and the current word representation $V\Ket{w_{i_j}}$ by concatenating the two representation vectors, multiplying by $W_{combine}$ and adding $b_{combine}$ before applying the $tanh(\cdot)$ function:
$$p_j = tanh(W_{combine}\Big[p_{j-1};V\Ket{w_{i_j}}\Big] + b_{combine})$$
Subsequently, each $p_j$ is passed along to the classifier, which applies the operation:
$$t_{i_j} = softmax(W_{classify}p_j + b_{classify})$$
to obtain the predicted tag $t_{i_j}$ for $w_{i_j}$.

@@

Training neural networks can be tricky since the backpropagation algorithm relies upon a learning rate parameter as well as initial weight and bias values. The learning rate should not be too large, otherwise the gradient descent will skip over local or global minima. However, it should also not be too small, otherwise training may take too long. It perhaps should vary over the course of training and decrease in order to better fine tune the fit to the minima. However, it is not obvious exactly which set of choices will be best.

@@

\subsubsection{AdaGrad}
We used the AdaGrad online learning method to train our model (Socher et al., 2013). Let $\theta$ be a vector of all parameters needed to train an RNN model and $\theta_t$ means the parameter vector before training on the $t$-th sentence. And let $g_{t,i}$ to be the gradient value of $i$-th parameter on the $t$-th sentence. Then to perform AdaGrad is just to update the parameter vector by
$$\theta_{{t+1},i} = \theta_{{t},i}  - \frac{\alpha}{\sqrt{\sum_{s=1}^t \theta_{{s},i}^2 }} g_{t,i}$$,
where the $\alpha$ is a constant to set the scale of the learning rate.

We compared the AdaGrad with fixed learning rate online training. The Figure \ref{Adagrad_fig} is the accuracy curve on training iteration. From Figure \ref{Adagrad_fig}, it is quite clear that the AdaGrad converges faster and better than fixed learning rate. The discrepancy between the initial positions is due to plotting after the first iteration over the training set, at which point we've already performed one AdaGrad update for each sentence.

Additionally, we found that performing an AdaGrad update after each word evaluation performed worse than updating once after each sentence evaluation.

@@

\begin{itemize}
\item \textbf{Same word test}:  We chose some words with more than one common tag and saw whether our model predicted those tags correctly. We tested on `to' (possible tags are TO and IN) and `work' (common possible tags are NN, VB and VBP). Our RNN predicts `to' with 0.80 in-domain and 0.77 out-of-domain accuracy, covering TO and IN. The RNN predicts `work' with 0.88 in-domain and 0.83 out-of-domain accuracy, covering NN, VB and VBP.
\item \textbf{Disconnected core module test}: We disconnected the sequential information passing by ignoring the input embeddings from lower layers, and backward gradient from upper layers. So although it is still an RNN, there is actually no information passed between layers. We trained this non-sequential RNN and the original RNN with 100 sentences and 100 iterations, same random seed and training parameters. The Figure \ref{Noseq_fig} shows the out-of-domain accuracy of two models by the number of iterations, which can be also seen from Figure \ref{Leftsize_fig}. The difference is not small. So we think the sequential information was passed and used at least partially correctly.
\end{itemize}
\begin{figure}
\includegraphics[scale=0.5]{outdomain_noseq.png}
\caption{Comparison of non-sequential RNN and original RNN for disconnected core module test}\label{Noseq_fig}
\end{figure}
